\documentclass[12pt]{article}

\usepackage{graphics}
\usepackage{epsfig}
\usepackage{times}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[table,dvipsnames]{xcolor}
\usepackage{tcolorbox}
\usepackage{hyperref}

\usepackage{tabularx}
\usepackage{amsthm}
\usepackage{eqparbox}
\usepackage{relsize}
\usepackage{color}
\usepackage{listings}
\usepackage[export]{adjustbox}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{soul}
\usepackage{booktabs,rotating}
\usepackage{array}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[normalem]{ulem}
\usepackage{xcolor}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{listings}
\definecolor{mycolor}{RGB}{211, 211, 211} % Define the color 'mycolor' as blue
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\newtheorem{definition}{Definition}
\write18{}

\lstdefinestyle{XMLStyle}{
  language=XML,
  keywordstyle=\color{blue},
  numberstyle=\tiny\color{codegray},
  stringstyle=\color{codepurple},
  morekeywords={E,S,M,T,Failure,line},
  basicstyle=\tiny\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  frame=none,
  breaklines=true,
  aboveskip=10pt,
  belowskip=10pt,
  numbers=none,
  xleftmargin=10.5em
}
\newcommand{\FRAME}{{\sc Frame}\xspace}
\newcommand{\MotorEaseB}{{\bfseries\scshape MotorEase}\xspace}
\newcommand{\MotorEase}{{\sc MotorEase}\xspace}
\newcommand{\MotorCheck}{{\sc MotorCheck}\xspace}
\newcommand{\Miracle}{{\sc Miracle}\xspace}
\newcommand{\AidUI}{{\sc AidUI}\xspace}
\newcommand{\AidUIs}{{\sc AidUI's~}\xspace}
 
% \usepackage{comment}
% \excludecomment{figure}
% \let\endfigure\relax

% %%% Remove the next two lines if you want the figures and tables at their place    
% \usepackage[nolists,nomarkers]{endfloat}
% \renewcommand{\processdelayedfloats}{}

%%% custom commands

\newcommand{\code}[1] {{\smaller\texttt{#1}}}

\usepackage{tikz}

\newcommand*\emptycirc[1][black]{\tikz\draw[#1] (0,0) circle (1ex);} 
\newcommand*\halfcirc[1][black]{%
    \begin{tikzpicture}
        \draw[fill=#1] (0,0)-- (90:1ex) arc (90:270:1ex) -- cycle[#1] ;
        \draw[#1] (0,0) circle (1ex);
    \end{tikzpicture}
  }
\newcommand*\fullcirc[1][black]{\tikz\fill[#1] (0,0) circle (1ex);} 

\newcommand{\SAT}{{SAT}}
\newcommand{\PAT}{{PAT}}
\newcommand{\NLC}{\textsf{NLC}}
\newcommand{\FC}{\textsf{FC}}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{\small{\textbf{
			\node[shape=circle,fill=mycolor,draw=black, inner sep=0.75pt] (char) {\textcolor{black}{#1}};}}}}


\newcommand{\subsubsubsec}[1]{\paragraph{#1}\mbox{}\\}

\definecolor{cOrange}{rgb}{0.93, 0.35, 0.0}
\definecolor{cPurple}{rgb}{0.55, 0.0, 0.55}




% <http://psl.cs.columbia.edu/phdczar/proposal.html>:
%
% The standard departmental thesis proposal format is the following:
%        30 pages
%        12 point type
%        1 inch margins all around = 6.5   inch column
%        (Total:  30 * 6.5   = 195 page-inches)
%
% For letter-size paper: 8.5 in x 11 in
% Latex Origin is 1''/1'', so measurements are relative to this.

\topmargin      0.0in
\headheight     0.0in
\headsep        0.0in
\oddsidemargin  0.0in
\evensidemargin 0.0in
\textheight     9.0in
\textwidth      6.5in
\linespread {1}



\input{macros.tex}

\begin{document}
\pagestyle{plain}
\pagenumbering{roman}
% \maketitle

\begin{titlepage}
   \begin{center}
       \vspace*{1cm}

       \textbf{\LARGE Engineering Accessible Software}

       \vspace{2.5cm}
        \textit{Thesis proposal} 
        
       \vspace{2.5cm}

       {\bf Arun Krishna Vajjala}  \\
Department of Computer Science \\
George Mason University\\
Fairfax, VA 22030\\
 akrishn@gmu.edu \\

\vfill
            
\textbf{Committee}\\
Kevin Moran, University of Central Florida (Chair)\\
Brittany Johnson-Matthews, George Mason University \\
Andrian Marcus, George Mason University \\ 
Thomas LaToza, George Mason University \\
Vivian Motti, George Mason University \\
    
\vspace{0.8cm}
 
    
\date{\today}
\end{center}
\end{titlepage}

\pagebreak

\begin{abstract}
Software has become an integral part of daily life, spanning web applications, smartphone apps, and desktop systems. As reliance on technology grows, ensuring accessibility for users of all abilities is important. This is not only an ethical motivation but also a governmental requirement. Consequently, there's been a notable increase in efforts to make software more accessible, driving the need for innovative tools to assist developers in this endeavor.

This thesis explores advancements in software engineering research, particularly in the area of accessibility testing and developer tools. Leveraging state-of-the-art techniques like computer vision and machine learning, novel approaches are being developed to automatically detect accessibility issues within software interfaces. 

The core challenge lies in developing models capable of understanding UI designs while capturing their semantic meanings accurately. Addressing this challenge, the thesis proposes the development of tools to assist developers in enhancing software accessibility and tackling future problems requiring a deeper understanding of UI semantics.

Specifically, the thesis proposes a computer vision-based tool to identify motor-impairment accessibility issues in Android mobile apps and demonstrates its efficacy against state-of-the-art baselines. Additionally, a screen understanding tool leveraging computer vision techniques is proposed to aid in screen recognition and retrieval tasks, notably improving search efficiency.

The thesis outlines the background, main contributions, and future directions of the proposed research. It highlights the importance of addressing software accessibility, explores new techniques for detecting accessibility issues and failures, and lays out a research plan for further investigation and development in this domain.


\end{abstract}

\pagebreak
\tableofcontents
\pagebreak
\listoffigures
\listoftables
\newpage
\cleardoublepage
\pagenumbering{arabic}


% input each section here .. 
\input{Sections/Introduction}
\input{Sections/Background}
\input{Sections/Thesis}
\input{Sections/Detecting}
% \input{Sections/FlakeFlaggerStudy}
% \input{Sections/FlakeFlaggerClassifier}
\input{Sections/Living}
% \input{Sections/FailureLogsStudy}
% \input{Sections/FailureLogsApproach}
\input{Sections/Categorizing}
\input{Sections/SLR}
% \input{Sections/ProposedWork}
\input{Sections/ResearchPlan}





% \subsection{Flaky Tests and Their Impacts}
% \subsection{Techniques in Detecting Flaky Tests}

% \section{Detecting Flaky Tests Without Rerun}
% \subsection{Rerun Study}
% \subsection{Flaky Tests Classifier}
% \subsubsection{Features Collections}
% \subsubsection{Dataset}
% \subsubsection{Evaluation of Flaky Tests Classifier}

% \section{Detecting Flaky Failures Using Logs}
% \subsection{Study Part}
% \subsection{Approach Part}
% \subsection{Evaluation of Detecting Using Failure Logs}

% \section{Proposed Work}
% \subsection{NA}
% \begin{itemize}
%     \item start with our already collected dataset ... 
%     \item How many s a test flake with different failures. 
%     \item OD NOD ( maybe both)
%     \item Root causes based on the logs.. 
%     \item fix a bug may eliminate the flakiness cause.. 
%     \item 
% \end{itemize}

% \section{Research Plan}




\newpage


\begin{footnotesize}
\bibliographystyle{plain}
% \bibliography{string,itu,rfc,i-d}
\bibliography{ProposalBib.bib}
\end{footnotesize}

\newpage



\end{document}


